Weight Lifting Error Prediction
========================================================
### Introduction
This is a project homework of **Practical Machine Learning** by Johns Hopkins on coursera.

### Data 

The training data for this project are available here: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv]

The test data are available here: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv]

The data for this project come from this source: [http://groupware.les.inf.puc-rio.br/har]. 

### Goal
The data is from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
My goal is using the machine learning techniques I learned to train a model based on the training data to predict test data.


### Method Selection
This is clearly a classification problem. 
```{r cache=TRUE}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

```{r echo=FALSE}
set.seed(33833)
library(caret)
```
```{r}
dim(training)
```
Based on the fact there're a lot of variables in the dataset, my first hunch is *randome forest*.

### Data Processing
There're a lot of rows containing all *NA* in the *testing* dataset, so I decided to exclude these variables from *training*
dataset.
```{r}
test<-testing[,-which(apply(testing,2,function(x)all(is.na(x))))]
```
Variables like X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window do not likely to be useful to predict the result, so I remove them too.
```{r}
select <- colnames(subset(test, select = -c(X, user_name, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp, new_window, problem_id)))
```

I split the training dataset into two parts, *train* for training the model, *mytest* for testing my model.
```{r}
inTrain <- createDataPartition(y=training$classe, p = 0.60, list=F)
training <- training[,c(select, "classe")]
train <- training[inTrain,]
mytest <- training[-inTrain,]
```

### Model Training
By default, *train()* use bootstrap method which is expensive for my laptop, so change it to *cross validation*
```{r cache=TRUE}
trControl = trainControl(method = "cv", number = 2)
modFit <- train(classe ~ .,data=train, method="rf", trControl = trControl)
```

### Cross Validation
```{r}
pred <- predict(modFit, mytest)
confusionMatrix(mytest$classe, pred)
```
The **Accuracy** is pretty high, based on the fact that **out of sample error** is 1 - accuracy, we can say the out of sample error is very low. So this model is good enough.

### Generating Results Files For Course Project Submission
```{r}
finalpred <- predict(modFit, test)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(as.character(finalpred))
```
